---
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MultiSynth Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
library(kableExtra)
```


# `augsynth`: Estimating treatment effects with staggered adoption

### The data

To show the features of the `multisynth` function we will use data on the effects of states implementing mandatory collective bargaining agreements for public sector unions [(Paglayan, 2018)](https://onlinelibrary.wiley.com/doi/full/10.1111/ajps.12388)

```{r results="hide", warning=F, message=F}
library(magrittr)
library(dplyr)
#devtools::install_github("ebenmichael/augsynth")
library(augsynth)
```

```{r }
data <- read.csv("https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/WGWMAV/3UHTLP", sep="\t")
```

The dataset contains several important variables that we'll use:

- `year`, `State`: The state and year of the measurement
- `YearCBrequired`: The year that the state adopted mandatory collective bargaining
- `lnppexpend`: Log per pupil expenditures in constant 2010 $

To run `multisynth`, we need to include a treatment status column that indicates which state is treated in a given year, we call this `cbr` below. We also restrict to the years 1959-1997 where we have yearly measurements of expenditures and drop Washington D.C. and Wisconsin from the analysis.

```{r }
analysis_df <-  data %>%
  filter(!State %in% c("DC", "WI"),
           year >= 1959, year <= 1997) %>%
  mutate(YearCBrequired = ifelse(is.na(YearCBrequired), 
                                   Inf, YearCBrequired),
           cbr = 1 * (year >= YearCBrequired))
```

## Partially pooled SCM with an intercept

To fit partially pooled synthetic controls, we need to give `multisynth` a formula of the form `outcome ~ treatment`, point it to the unit and time variables, and choose the level of partial pooling `nu`. Setting `nu = 0` fits a separate synthetic control for each treated unit and setting `nu = 1` fits fully pooled synthetic controls. If we don't set `nu`, `multisynth` will choose a heuristic value based on how well separate synthetic controls balance the overall average.
By default, `multisynth` includes an intercept shift along with the weights; we can exclude the intercept shift by setting `fixedeff = F`.
We can also set the number of pre-treatment time periods (lags) that we want to balance with the `n_lags` argument and the number of post-treatment time periods (leads) that we want to estimate with the `n_leads` argument. By default `multisynth` sets `n_lags` and `n_leads` to the number of pre-treatment and post-treatment periods for the last treated unit, respectively.

```{r }
# with a choice of nu
ppool_syn <- multisynth(lnppexpend ~ cbr, State, year, 
                        nu = 0.5, analysis_df)
# with default nu
ppool_syn <- multisynth(lnppexpend ~ cbr, State, year, 
                        analysis_df)
print(ppool_syn$nu)
ppool_syn
```

Using the `summary` function, we'll compute the treatment effects and standard errors and confidence intervals for all treated units as well as the average via the wild bootstrap. (This takes a bit of time so we'll store the output) We can also change the significant level associated with the confidence intervals by setting the `alpha` argument, by default `alpha = 0.05`.

```{r}
ppool_syn_summ <- summary(ppool_syn)
```

We can then report the level of global and individual balance as well as estimates for the average.

```{r }
ppool_syn_summ
```

`nopool_syn_summ$att` is a dataframe that contains all of the point estimates, standard errors, and lower/upper confidence limits. `Time = NA` denotes the effect averaged across the post treatment periods.

```{r echo = F}
ppool_syn_summ$att %>%
  filter(Time >= 0) %>%
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options =c("hover", "responsive"))
```

We can also visually display both the pre-treatment balance and the estimated treatment effects.

```{r ppool_syn_plot, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_summ)
```

And again we can hone in on the average effects.

```{r ppool_syn_plot_avg, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_summ, levels = "Average")
```


### Collapsing into time cohorts

We can also collapse treated units with the same treatment time into _time cohorts_, and find one synthetic control per time cohort by setting `time_cohort = TRUE`. When the number of distinct treatment times is much smaller than the number of treated units, this will run significantly faster.

```{r }
# with default nu
ppool_syn_time <- multisynth(lnppexpend ~ cbr, State, year,
                        analysis_df, time_cohort = TRUE)
print(ppool_syn_time$nu)
ppool_syn_time
```

We can then compute effects for the overall average as well as for each treatment time cohort, rather than individual units.

```{r}
ppool_syn_time_summ <- summary(ppool_syn_time)
ppool_syn_time_summ
```

```{r echo = F}
ppool_syn_time_summ$att %>%
  filter(Time >= 0) %>%
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options =c("hover", "responsive"))
```

Again we can plot the effects.

```{r ppool_syn_time_plot, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_time_summ)
```


### Including auxiliary covariates

We can also include an additional set of covariates to balance along with the pre-treatment outcomes. First, let's create a data frame with the values of some covariates in a few different years:

```{r cov_data}
data %>%
  select(State, year, agr, pnwht, purban, perinc, studteachratio) %>%
  group_by(State) %>%
  summarise(perinc_1959 = perinc[year == 1959],
            studteachratio_1959 = studteachratio[year == 1959]) %>% 
  # filter to lower 48 where we have data
  filter(!State %in% c("AK", "HI"))  -> cov_data
analysis_df %>%
  inner_join(cov_data, by = "State") -> analysis_df_covs
```

To include auxiliary covariates, we can add them in to the formula after `|`. This will balance the auxiliary covariates along with the pre-treatment outcomes simultanouesly. If the covariates vary during the pre-treatment periods, `multisynth` will use the average pre-treatment value. We can change this behavior by including our own custom aggregation function via the `cov_agg` argument.
```{r cov_syn}
# with default nu
ppool_syn_cov <- multisynth(lnppexpend ~ cbr | perinc_1959 + studteachratio_1959,
                            State, year, analysis_df_covs)
print(ppool_syn_cov$nu)
ppool_syn_cov
```

Again we can compute effects, along with their standard errors and confidence intervals, and plot.
```{r}
ppool_syn_cov_summ <- summary(ppool_syn_cov)
ppool_syn_cov_summ
```

```{r echo = F}
ppool_syn_cov_summ$att %>%
  filter(Time >= 0) %>%
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options =c("hover", "responsive"))
```

Again we can plot the effects.
```{r ppool_syn_cov_plot, fig.width=8, fig.height=4.5, fig.align="center", warning=F, message=F}
plot(ppool_syn_cov_summ, levels = "Average")
```