---
output: github_document
---

```{r setup, include = F}
# devtools::install_github("Hemken/Statamarkdown")
library(Statamarkdown)
```

# Lalonde

This dataset is one of the most commonly used dataset in econometrics based on [Lalonde (1986)](https://econpapers.repec.org/article/aeaaecrev/v_3a76_3ay_3a1986_3ai_3a4_3ap_3a604-20.htm) and [Dehejia and Wahba (2002)](https://www.uh.edu/~adkugler/Dehejia&Wahba.pdf). Both the paper by Lalonde and Dehejia and Wahba both wanted to evaluate causal inference methods using non-experimental data. Cleverly, they start with an experimental dataset to estimate the 'true' causal effect and then use non-experimental data to evaluate an econometric method.

In [Causal Inference 1](github.com/Mixtape-Sessions/Causal-Inference-1), we have already shown that the nonexperimental dataset performs poorly at estimating the effects of a job training program on earnings, finding a substantial negative effect. That lab shows that matching-based methods are effective at recovering effects close to the experimental effect. This lab will study difference-in-differences estimators of the effects.

1. We will first perform analysis on the experimental dataset `https://raw.github.com/Mixtape-Sessions/Causal-Inference-2/master/Labs/Lalonde/lalonde_exp_panel.dta`

```{r load-exp-data}
library(tidyverse)
library(fixest)
library(DRDID) # devtools::install_github("pedrohcgs/DRDID")
library(haven)

# 1. Experimental data
df_exp <- haven::read_dta("https://raw.github.com/Mixtape-Sessions/Causal-Inference-2/master/Lab/Lalonde/lalonde_exp_panel.dta")
```

a. As a reminder, perform a simple difference-in-means on the experimental dataset to estimate the "treatment effect". Use only the year `78` and compare average real earnings for the treated and the control groups. This is our baseline estimate of the treatment effect

```{r difference-in-means-exp}
# ---- Difference-in-means
feols(
  re ~ i(treat),
  data = df_exp[df_exp$year == 78, ], vcov = "hc1"
)
```

b. Estimate the effect of treatment, `treat`, on real-earnings, `re` in a difference-in-differences estimator. This will be the "true" treatment effect estimate that we will try to recreate with the non-experimental CPS sample. 

```{r diff-in-diff-exp}
# ---- Difference-in-Differences Estimate
feols(
  re ~ i(treat) | id + year, 
  data = df_exp, vcov = "hc1"
)
```

2. Now, we turn to the non-experimental dataset `https://raw.github.com/Mixtape-Sessions/Causal-Inference-2/master/Labs/Lalonde/lalonde_nonexp_panel.dta`. 

```{r load-nonexp-data}
# 2. CPS data
df_nonexp <- haven::read_dta("https://raw.github.com/Mixtape-Sessions/Causal-Inference-2/master/Lab/Lalonde/lalonde_nonexp_panel.dta")
```

a. As a reminder, perform a simple difference-in-means on the combined dataset to estimate the "treatment effect". Use only the year `78` and compare average real earnings for the treated and the control groups.

```{r difference-in-means-nonexp}
# ---- Difference-in-means
feols(
  re ~ i(treat),
  data = df_nonexp[df_nonexp$year == 78, ], vcov = "hc1"
)
```

b. Estimate a difference-in-differences estimator for the non-experimental group.

```{r diff-in-diff-nonexp}
# ---- Initial Difference-in-difference
feols(
  re ~ i(treat) | id + year, 
  data = df_nonexp, vcov = "hc1"
)
```

c. Use the `DRDID` command to estimate a doubly-robust DID with covariates `age + agesq + agecube + educ + educsq + marr + nodegree + black + hisp`

```{r drdid}
# ---- Double-robust DID
drdid(
  yname = "re", tname = "year", idname = "id", dname = "ever_treated", 
  xformla = ~ age + agesq + agecube + educ + educsq +
    marr + nodegree + black + hisp + re74 + u74,
  data = df_nonexp %>% filter(year == 75 | year == 78)
)
```

